\Large\bf Scalable and Probabilistic Leaderless BFT Consensus through Metastability

### 2020-11-09 paper_fr.tex
Typos A.1 A.2
Analysis - A3

\paragraph{Attaques Sybil}
%O Consensus protocols provide their guarantees based on assumptions that only a fraction of participants are adversarial.
Les protocoles de consensus fournissent leurs garanties fondées sur la supposition que seule une fraction des participants sont malveillants.
%O These bounds could be violated if the network is naively left open to arbitrary participants.
Ces limites peuvent être violées si le réseau est naïvement laissé ouvert à des participants quelconques.
%O In particular, a Sybil attack~\cite{douceur2002sybil}, wherein a large number of identities are generated by an adversary, could be used to exceed the bounds.
En particulier, une attaque Sybil~\cite{douceur2002sybil}, où un grand nombre d'identités sont générées par un adversaire, peut être employée pour dépasser ces limites.

@@

%O \paragraph{Irreversibility}
\paragraph{Irreversibilité}
%We now discuss the core results of irreversibility properties of our family of protocols.
%O In Section~\ref{sec:analysis}, we discussed the loose Chvatal bound which provided intuitive understanding into the strong irreversibility dynamics of our core subsampling mechanism. In particular, once the network drifts to some majority value, it tends to revert back with only an exponentially small probability. We compute the closed-form expression for reversibility, and show that it is exponentially small.
En section ~\ref{sec:analysis}, nous avons traité de la borne \emph{loose} % HOULA
de Chtaval qui a fourni une compréhension intuitive de la dynamique de forte irréversibilité au cœur de notre mécanisme de sous-échantillonnage. En particulier, une fois que le réseau tend vers une valeur majoritaire donnée, il ne peut s'inverser qu'avec une probabilité exponentiellement petite. Nous calculons l'expression de forme fermée pour la réversibilité, et nous montrons qu'elle est exponentiellement petite.
\begin{theorem}
\label{theorem:slush_prob_convergence_minority}
%O Let $\xi_\delta$ be the probability of absorption into the all-red state ($s_0$), starting from a drift of $\delta$ (i.e. $\delta$ drift away from $n/2$). Then, assuming $\delta > 1$, 
Soit $\xi_\delta$ la probabilité d'absorption dans l'état totalement rouge ($s_0$), commençant par un drift de $\delta$ (c'est-à-dire que $\delta$ tendent à s'éloigner de $n/2$). Alors, en supposant $\delta > 1$,
\begin{equation}
\xi_\delta = 1 - \ddfrac{\sum_{l = 1}^{\delta} \prod_{i = 1}^{l-1} \mu_i^2 \prod_{j = l}^{n-l}\lambda_j}{2\sum_{l = 1}^{n/2}\prod_{i=1}^{l-1}\mu_i^2\prod_{j=l}^{n-l}\mu_j}
\end{equation}
et
% \begin{equation}
% \xi_\delta - \xi_{\delta+1} = \ddfrac{\mu_1^2\dots\mu_{\delta}^2\lambda_{\delta+1}\dots\lambda_{n-\delta}}{\Phi(\alpha, k, n)}
% \end{equation}
% where $\Phi(\alpha, k, n)$ is a constant dependent on $\alpha$, $k$, and $n$, but otherwise independent of starting configuration. 
\begin{equation}
\begin{split}
\ddfrac{\xi_{\delta} - \xi_{\delta+1}}{\xi_{\delta+1} - \xi_{\delta+2}} &= \mathcal{u}_{\delta+1} = \ddfrac{\lambda_{\delta+1}}{\mu_{\delta+1}} \\
&\approx \ddfrac{n-\delta-1 \sum_{j = \alpha}^{k}\ddfrac{(n-\delta-1)^k (\delta+1)^{k-j}}{n^{2k - j}}}{\delta+1 \sum_{j = \alpha}^{k}\ddfrac{(\delta+1)^k (n-\delta-1)^{k-j}}{n^{2k - j}}}
\end{split}
\end{equation}
%O where from now on we refer to $\mathcal{u}_{\delta+1}$ as the drift of the process. 
où à partir de maintenant nous nous référons à $\mathcal{u}_{\delta+1}$ par le drift du processus.
\end{theorem}

\begin{proof}
%O Our results are derived based on constructions from Tan~\cite{tan1976absorption}. We construct a sub-matrix of $Q$, denoted $B$, as shown in Figure~\ref{fig:matrixB}.
Nos résultats sont dérivés en se fondant sur les constructions de Tan~\cite{tan1976absorption}. Nous construisons une sous-matrice de $Q$, notée $B$, comme on le voit en figure~\ref{fig:matrixB}.
\begin{figure*}
\[B = 
\begin{bmatrix}
    -(\lambda_1 + \mu_1) & \lambda_1 & 0 & \cdots & \cdots & 0 \\
    \mu_2 & -(\lambda_2 + \mu_2) & \lambda_2 & 0 & \cdots & 0\\
    0 & \mu_3 & -(\lambda_3 + \mu_3) & \lambda_3 & \cdots & 0\\
    \vdots & \vdots & \ddots & \ddots & \ddots & \vdots\\
    \vdots & \vdots & \mu_{n-3} & -(\lambda_{n-2} + \mu_{n-2}) & \lambda_{n-3} & 0\\
    \vdots & \dots & 0 & \mu_{n-1} & -(\lambda_{n-2} + \mu_{n-2}) & \lambda_{n-2}\\
    0 & \dots & 0 & 0 & \mu_{n-1} & -(\lambda_{n-1} + \mu_{n-1})
\end{bmatrix}
\]
\caption{Matrice $B$.}
\label{fig:matrixB}
\end{figure*}
Let $W_1'$ = $(\mu_1, 0, \dots, 0)$, $W_{n-1}'$ = $(0, \dots, 0, \lambda_{n-1})$. Then, we can express $Q$ as 
\[
    Q =
    \begin{bmatrix}
        0 & \dots & 0\\
        W_1 & B & W_{n-1}\\
        0 & \dots & 0
    \end{bmatrix}
\]
%O As a reminder, the stationary distribution can be found via $\lim_{t \rightarrow \infty} P(t) = e^{Qt}$, where we have
Rappelons que la distribution stationnaire peut être trouvée \emph{via} $\lim_{t \rightarrow \infty} P(t) = e^{Qt}$, où nous avons
\[
    e^{Qt} = \sum_{i = 0}^{\infty} \frac{t^i}{i!} Q^i = \sum_{i = 0}^{\infty} \frac{t^i}{i!}
    \begin{bmatrix}
        0 & \dots & 0\\
        B^{i-1}W_1 & B^i & B^{i-1}W_{n-1}\\
        0 & \dots & 0
    \end{bmatrix}
\]
As Tan (eq. 2.3) shows, we have
Comme le montre Tan (eq. 2.3), nous avons
\[
    \xi(t) = B^{-1}\left[\sum_{i = 0}^{\infty} B^i  - \mathbb{I}_{n-1} \right] W_1
\]
%O Since we want the ultimate probabilities, we have that 
Comme nous voulons les probabilités ultimes, nous avons 
\[
    \xi = \lim_{t \rightarrow \infty} \xi(t) = -B^{-1}W_1
\]
%O We can explicitly compute $\xi_\delta$ in terms of our rates $\mu_i$ and $\lambda_i$, getting 
Nous pouvons explicitement calculer $\xi_\delta$ en termes de nos taux $\mu_i$ et $\lambda_i$, obtenant ainsi
\[
    \xi_\delta = \ddfrac{\sum_{l = 1}^{n-\delta}\prod_{i = 1}^{n-l}\mu_i \prod_{j = n-l+1}^{n-1}\lambda_j}{\sum_{l = 1}^{n}\prod_{i = 1}^{n-l}\mu_i \prod_{j = n-l+1}^{n-1}\lambda_j}
\]
%O However, we note that $u_{i} = \lambda_{n-i}$. Algebraic manipulation from this observation leads to the two equations in the theorem. This expression is strictly lower than the Chvatal bounds used in Section~\ref{sec:analysis}.
Cependant, notons que $u_{i} = \lambda_{n-i}$. La manipulation algébrique à partir de cette observation mène aux deux équations du théorème. Cette expression est strictement inférieure aux bornes de Chtaval utilisées en section~\ref{sec:analysis}.
\end{proof}

@@

%O Theorem~\ref{theorem:mean-convergence-time} leads to the next lemma that captures property P2, under the assumption that at the beginning of the protocol, one proposal has at least $\alpha$ support in the network. 
Le théorème~\ref{theorem:mean-convergence-time} mène au lemme suivant qui capture la propriété P2, en suppposant qu'au début du protocole, une proposition a un support d'au moins $\alpha$ dans le réseau.
\begin{lemma}
%O Slush reaches an absorbing state in finite time almost surely.
Slush atteint presque sûrement un état absorbant dans un temps fini.
\label{lemma:finitetermination}
\end{lemma}

@@

\subsection{Snowflake}
\label{subsection:appendix_snowflake}
In Snowflake, the sampled set of nodes includes Byzantine nodes.
We introduce the decision function $\mathcal{D}(*)$, which is constructed by having each node also keep track of the total number of consecutive times it has sampled a majority of the same color ($\beta$). 
Finally, we introduce a function called $\mathcal{A}(\mathcal{S}_t)$, the adversarial strategy, that takes as parameters the entire configuration of the network at time $t$, as well as the next set of nodes chosen by the scheduler to execute, and as a side-effect, modifies the set of nodes $\mathcal{B}$ to some arbitrary configuration of colors.

In order for our prior framework to apply to Snowflake, we must deal with a key subtlety. 
Unlike in Slush, where it is clear that once the network has reached one of the converging states and therefore may not revert back, this no longer applies to Snowflake, since any adversary $f \geq \alpha$ has strictly positive probability of reverting the system, albeit this probability may be infinitesimally small. 
The CTMC is flexible enough to deal with a system where there is only one absorbing state, but the long-term behavior of the system is no longer meaningful since, after an infinite amount of time, the system is guaranteed to revert, violating safety. 
We could trivially bound the amount of time, and show safety using this bounded time assumption by simply characterizing the distribution of $e^{tQ}$, where $Q$ is the generator. 
However, we can make the following observation: if the probability of going from state $c$ (all-blue) to $c-1$ is exponentially small, then it will take the attacker exponential time (in expectation; note, this is a lower bound, and in reality it will take much longer) to succeed in reverting the system. 
Hence, we can assume that once all correct nodes are the same color, the attack from the adversary will terminate since it is impractical to continue an attack. 
In fact, under reasonably bounded timeframes, the variational distance between the exact approach and the approximation is very small. 
We leave details to the accompanying paper, but we briefly discuss how analysis proceeds for Snowflake.

As stated in Section~\ref{sec:analysis}, the way to analyze the adversary using the same construction as in Slush is to condition reversibility on the first node $u$ deciding on blue, which can happen at any state (as specified by $\mathcal{D}(*)$). At that point, the adversarial strategy collapses to a single function, which is to continually vote for red. The probabilities of reversibility, for all states $\{1, \dots, c-1\}$ must encode the probability that additional blue nodes commit, and the single function of the adversary. The birth and death rates are transformed as follows:
\[
    \begin{cases}
        \mu_i = &i(1 - \mathbb{I}[\mathcal{D}(*, i, \mathbb{B})])\ \mathcal{H}(\mathcal{N}, c-i + f, k, \alpha)\\
        \lambda_i = &(c-i)(1 - \mathbb{I}[\mathcal{D}(*, c-i, \mathbb{R})])\ \mathcal{H}(\mathcal{N}, i, k, \alpha)\\
    \end{cases}
\]
From here on, the analysis is the same as in Slush. Under various $k$ and $\beta$, we can find the minimal $\alpha$ that provides the system strong irreversibility properties.

The next lemma captures P3, and the proof follows from central limit theorem. 
\begin{lemma}
If $f < \Oh{\sqrt{n}}$, and $\alpha = \floor{k/2} + 1$, then Snowflake terminates in $\Oh{\log n}$ rounds with high probability. 
\label{lemma:centrallimit}
\end{lemma}
\begin{proof}
The results follows from central limit theorem, wherein for $\alpha = \floor{k/2} + 1$, the expected bias in the network after sampling will be $\Oh{\sqrt{n}}$. An adversary smaller than this bias will be unable to keep the network in a fully-bivalent state for more than a constant number of rounds. The logarithmic factor remains from the mixing time lower bound. 
\end{proof}

>>>>

\paragraph{Attaques Sybil}
%O Consensus protocols provide their guarantees based on assumptions that only a fraction of participants are adversarial.
Les protocoles de consensus fournissent leurs garanties fondées sur la supposition que seule une fraction des participants sont antagonistes.
%O These bounds could be violated if the network is naively left open to arbitrary participants.
Ces limites peuvent être violées si le réseau est naïvement laissé ouvert à des participants quelconques.
%O In particular, a Sybil attack~\cite{douceur2002sybil}, wherein a large number of identities are generated by an adversary, could be used to exceed the bounds.
En particulier, une attaque Sybil~\cite{douceur2002sybil}, où un grand nombre d'identités sont générées par un adversaire, peut être employée pour dépasser ces limites.

@@

%O \paragraph{Irreversibility}
\paragraph{Irréversibilité}
%We now discuss the core results of irreversibility properties of our family of protocols.
%O In Section~\ref{sec:analysis}, we discussed the loose Chvatal bound which provided intuitive understanding into the strong irreversibility dynamics of our core subsampling mechanism. In particular, once the network drifts to some majority value, it tends to revert back with only an exponentially small probability. We compute the closed-form expression for reversibility, and show that it is exponentially small.
En section ~\ref{sec:analysis}, nous avons traité de la borne \emph{loose} % HOULA
de Chtaval qui a fourni une compréhension intuitive de la dynamique de forte irréversibilité au cœur de notre mécanisme de sous-échantillonnage. En particulier, une fois que le réseau tend vers une valeur majoritaire donnée, il ne peut s'inverser qu'avec une probabilité exponentiellement petite. Nous calculons l'expression de forme fermée pour la réversibilité, et nous montrons qu'elle est exponentiellement petite.
\begin{theorem}
\label{theorem:slush_prob_convergence_minority}
%O Let $\xi_\delta$ be the probability of absorption into the all-red state ($s_0$), starting from a drift of $\delta$ (i.e. $\delta$ drift away from $n/2$). Then, assuming $\delta > 1$, 
Soit $\xi_\delta$ la probabilité d'absorption dans l'état totalement rouge ($s_0$), commençant par un drift de $\delta$ (c'est-à-dire que $\delta$ tendent à s'éloigner de $n/2$). Alors, en supposant $\delta > 1$,
\begin{equation}
\xi_\delta = 1 - \ddfrac{\sum_{l = 1}^{\delta} \prod_{i = 1}^{l-1} \mu_i^2 \prod_{j = l}^{n-l}\lambda_j}{2\sum_{l = 1}^{n/2}\prod_{i=1}^{l-1}\mu_i^2\prod_{j=l}^{n-l}\mu_j}
\end{equation}
et
% \begin{equation}
% \xi_\delta - \xi_{\delta+1} = \ddfrac{\mu_1^2\dots\mu_{\delta}^2\lambda_{\delta+1}\dots\lambda_{n-\delta}}{\Phi(\alpha, k, n)}
% \end{equation}
% where $\Phi(\alpha, k, n)$ is a constant dependent on $\alpha$, $k$, and $n$, but otherwise independent of starting configuration. 
\begin{equation}
\begin{split}
\ddfrac{\xi_{\delta} - \xi_{\delta+1}}{\xi_{\delta+1} - \xi_{\delta+2}} &= \mathcal{u}_{\delta+1} = \ddfrac{\lambda_{\delta+1}}{\mu_{\delta+1}} \\
&\approx \ddfrac{n-\delta-1 \sum_{j = \alpha}^{k}\ddfrac{(n-\delta-1)^k (\delta+1)^{k-j}}{n^{2k - j}}}{\delta+1 \sum_{j = \alpha}^{k}\ddfrac{(\delta+1)^k (n-\delta-1)^{k-j}}{n^{2k - j}}}
\end{split}
\end{equation}
%O where from now on we refer to $\mathcal{u}_{\delta+1}$ as the drift of the process. 
où à partir de maintenant nous nous référons à $\mathcal{u}_{\delta+1}$ par le drift du processus.
\end{theorem}

\begin{proof}
%O Our results are derived based on constructions from Tan~\cite{tan1976absorption}. We construct a sub-matrix of $Q$, denoted $B$, as shown in Figure~\ref{fig:matrixB}.
Nos résultats sont dérivés en se fondant sur les constructions de Tan~\cite{tan1976absorption}. Nous construisons une sous-matrice de $Q$, notée $B$, comme on le voit en figure~\ref{fig:matrixB}.
\begin{figure*}
\[B = 
\begin{bmatrix}
    -(\lambda_1 + \mu_1) & \lambda_1 & 0 & \cdots & \cdots & 0 \\
    \mu_2 & -(\lambda_2 + \mu_2) & \lambda_2 & 0 & \cdots & 0\\
    0 & \mu_3 & -(\lambda_3 + \mu_3) & \lambda_3 & \cdots & 0\\
    \vdots & \vdots & \ddots & \ddots & \ddots & \vdots\\
    \vdots & \vdots & \mu_{n-3} & -(\lambda_{n-2} + \mu_{n-2}) & \lambda_{n-3} & 0\\
    \vdots & \dots & 0 & \mu_{n-1} & -(\lambda_{n-2} + \mu_{n-2}) & \lambda_{n-2}\\
    0 & \dots & 0 & 0 & \mu_{n-1} & -(\lambda_{n-1} + \mu_{n-1})
\end{bmatrix}
\]
\caption{Matrice $B$.}
\label{fig:matrixB}
\end{figure*}
Let $W_1'$ = $(\mu_1, 0, \dots, 0)$, $W_{n-1}'$ = $(0, \dots, 0, \lambda_{n-1})$. Then, we can express $Q$ as 
\[
    Q =
    \begin{bmatrix}
        0 & \dots & 0\\
        W_1 & B & W_{n-1}\\
        0 & \dots & 0
    \end{bmatrix}
\]
%O As a reminder, the stationary distribution can be found via $\lim_{t \rightarrow \infty} P(t) = e^{Qt}$, where we have
Rappelons que la distribution stationnaire peut être trouvée \emph{via} $\lim_{t \rightarrow \infty} P(t) = e^{Qt}$, où nous avons
\[
    e^{Qt} = \sum_{i = 0}^{\infty} \frac{t^i}{i!} Q^i = \sum_{i = 0}^{\infty} \frac{t^i}{i!}
    \begin{bmatrix}
        0 & \dots & 0\\
        B^{i-1}W_1 & B^i & B^{i-1}W_{n-1}\\
        0 & \dots & 0
    \end{bmatrix}
\]
%O As Tan (eq. 2.3) shows, we have
Comme le montre Tan (eq. 2.3), nous avons
\[
    \xi(t) = B^{-1}\left[\sum_{i = 0}^{\infty} B^i  - \mathbb{I}_{n-1} \right] W_1
\]
%O Since we want the ultimate probabilities, we have that 
Comme nous voulons les probabilités ultimes, nous avons 
\[
    \xi = \lim_{t \rightarrow \infty} \xi(t) = -B^{-1}W_1
\]
%O We can explicitly compute $\xi_\delta$ in terms of our rates $\mu_i$ and $\lambda_i$, getting 
Nous pouvons explicitement calculer $\xi_\delta$ en termes de nos taux $\mu_i$ et $\lambda_i$, obtenant ainsi
\[
    \xi_\delta = \ddfrac{\sum_{l = 1}^{n-\delta}\prod_{i = 1}^{n-l}\mu_i \prod_{j = n-l+1}^{n-1}\lambda_j}{\sum_{l = 1}^{n}\prod_{i = 1}^{n-l}\mu_i \prod_{j = n-l+1}^{n-1}\lambda_j}
\]
%O However, we note that $u_{i} = \lambda_{n-i}$. Algebraic manipulation from this observation leads to the two equations in the theorem. This expression is strictly lower than the Chvatal bounds used in Section~\ref{sec:analysis}.
Cependant, notons que $u_{i} = \lambda_{n-i}$. La manipulation algébrique à partir de cette observation mène aux deux équations du théorème. Cette expression est strictement inférieure aux bornes de Chtaval utilisées en section~\ref{sec:analysis}.
\end{proof}

@@

%O Theorem~\ref{theorem:mean-convergence-time} leads to the next lemma that captures property P2, under the assumption that at the beginning of the protocol, one proposal has at least $\alpha$ support in the network. 
Le théorème~\ref{theorem:mean-convergence-time} mène au lemme suivant qui reflète % HOULA
la propriété P2, en suppposant qu'au début du protocole, une proposition possède un support d'au moins $\alpha$ dans le réseau.
\begin{lemma}
%O Slush reaches an absorbing state in finite time almost surely.
Slush atteint presque sûrement un état absorbant dans un temps fini.
\label{lemma:finitetermination}
\end{lemma}

@@

\subsection{Snowflake}
\label{subsection:appendix_snowflake}
%O In Snowflake, the sampled set of nodes includes Byzantine nodes.
Dans Snowflake, l'ensemble des nœuds échantillonnés comprend des nœuds byzantins.
%O We introduce the decision function $\mathcal{D}(*)$, which is constructed by having each node also keep track of the total number of consecutive times it has sampled a majority of the same color ($\beta$). 
Nous introduisons finalement une fonction appelée $\mathcal{A}(\mathcal{S}_t)$ qui est construite en faisant conserver à chaque nœud la trace du nombre total de fois consécutives qu'il a échantillonné une majorité de la couleur d'échantillon ($\beta$).
%O Finally, we introduce a function called $\mathcal{A}(\mathcal{S}_t)$, the adversarial strategy, that takes as parameters the entire configuration of the network at time $t$, as well as the next set of nodes chosen by the scheduler to execute, and as a side-effect, modifies the set of nodes $\mathcal{B}$ to some arbitrary configuration of colors.
Enfin, nous introduisons une fonction appelée $\mathcal{A}(\mathcal{S}_t)$, la stratégie contradictoire, qui prend comme paramètres toute la configuration du réseau à l'instant $t$, ainsi que l'ensemble suivant de nœuds choisi par l'ordonnanceur pour exécution, avec pour effet secondaire la modification de l'ensemble de nœuds $\mathcal{B}$ vers une configuration arbitraire de couleurs.

%O In order for our prior framework to apply to Snowflake, we must deal with a key subtlety. 
Afin d'appliquer notre framework antérieur à Snowflake, nous devons porter notre attention sur une subtilité essentielle.
%O Unlike in Slush, where it is clear that once the network has reached one of the converging states and therefore may not revert back, this no longer applies to Snowflake, since any adversary $f \geq \alpha$ has strictly positive probability of reverting the system, albeit this probability may be infinitesimally small. 
Contrairement à Slush, où il est clair qu'une fois que le réseau a atteint l'un des états convergents et ne reviendra donc pas en arrière, cela ne s'applique pas à Snowflake, puisque tout adversaire $f \geq \alpha$ a une probabilité strictement positive d'inverser la tendance du système, fût-elle infinitésimale.
%O The CTMC is flexible enough to deal with a system where there is only one absorbing state, but the long-term behavior of the system is no longer meaningful since, after an infinite amount of time, the system is guaranteed to revert, violating safety. 
Le processus de Markov à temps continu est assez souple pour s'accommoder d'un système où il n'existe qu'un seul état absorbant, mais le comportement du système sur le long terme n'est plus significatif puisque, après un laps de temps infini, le système doit s'inverser, violant ainsi la sécurité de fonctionnement.
%O We could trivially bound the amount of time, and show safety using this bounded time assumption by simply characterizing the distribution of $e^{tQ}$, where $Q$ is the generator. 
Nous pouvons borner trivialement l'intervalle de temps et montrer la sûreté de fonctionnement grâce à ce bornage en nous contentant de caractériser la distribution de $e^{tQ}$, où $Q$ est le générateur.
%O However, we can make the following observation: if the probability of going from state $c$ (all-blue) to $c-1$ is exponentially small, then it will take the attacker exponential time (in expectation; note, this is a lower bound, and in reality it will take much longer) to succeed in reverting the system. 
Nous pouvons cependant faire l'observation suivante : si la probabilité d'aller de l'état $c$ (totalement bleu) à $c-1$ est exponentiellement petite, alors il faudra à l'attaquant un temps exponentiel (attendu\,; on note que c'est une borne inférieure et, en réalité, il faudra beaucoup plus longtemps) pour réussir à inverser l'état du système.
%O Hence, we can assume that once all correct nodes are the same color, the attack from the adversary will terminate since it is impractical to continue an attack. 
Nous pouvons donc suppposer qu'une fois que tous les nœuds corrects sont de la même couleur, l'attaque de l'adversaire se termine puisqu'il serait irréaliste de la poursuivre.

%O In fact, under reasonably bounded timeframes, the variational distance between the exact approach and the approximation is very small. 
En fait, dans un cadre temporel raisonnablement borné, la distance variationnelle entre l'approche exacte et l'approximation est très petite.
%O We leave details to the accompanying paper, but we briefly discuss how analysis proceeds for Snowflake. 
Nous laissons les détails à l'article qui accompagne celui-ci, mais nous discutons brièvement du déroulement de l'analyse pour Snowflake.

%O As stated in Section~\ref{sec:analysis}, the way to analyze the adversary using the same construction as in Slush is to condition reversibility on the first node $u$ deciding on blue, which can happen at any state (as specified by $\mathcal{D}(*)$). At that point, the adversarial strategy collapses to a single function, which is to continually vote for red. The probabilities of reversibility, for all states $\{1, \dots, c-1\}$ must encode the probability that additional blue nodes commit, and the single function of the adversary. The birth and death rates are transformed as follows:
Comme il en a été fait mention en section~\ref{sec:analysis}, la manière d'analyser l'adversaire en employant la même construction que dans Slush consiste à conditionner la réversibilité sur le premier nœud $u$ se décidant pour le bleu, ce qui peut se produire à n'importe quel état (comme le spécifie $\mathcal{D}(*)$). À ce point, la stratégie contradictoire se réduit à une seule fonction, qui est de continuellement voter pour rouge. Les probabilités de réversibilité, pour tous les états $\{1, \dots, c-1\}$ doivent coder la probabilité que des nœuds bleus s'engagent ansi que la fonction unique de l'adversaire. Les taux de naissance et de mort sont transformés comme suit\,:
\[
    \begin{cases}
        \mu_i = &i(1 - \mathbb{I}[\mathcal{D}(*, i, \mathbb{B})])\ \mathcal{H}(\mathcal{N}, c-i + f, k, \alpha)\\
        \lambda_i = &(c-i)(1 - \mathbb{I}[\mathcal{D}(*, c-i, \mathbb{R})])\ \mathcal{H}(\mathcal{N}, i, k, \alpha)\\
    \end{cases}
\]
%O From here on, the analysis is the same as in Slush. Under various $k$ and $\beta$, we can find the minimal $\alpha$ that provides the system strong irreversibility properties. 
À partir de maintenant, l'analyse est la même que pour Slush. Pour divers $k$ et $\beta$, nous pouvons trouver le $\alpha$ minimal qui fournit au système des propriétés de forte irréversibilité.

%O The next lemma captures P3, and the proof follows from central limit theorem. 
Le lemme suivant reflète P3 et la preuve procède du théorème central limite.
\begin{lemma}
%O If $f < \Oh{\sqrt{n}}$, and $\alpha = \floor{k/2} + 1$, then Snowflake terminates in $\Oh{\log n}$ rounds with high probability. 
Si $f < \Oh{\sqrt{n}}$, et $\alpha = \floor{k/2} + 1$, alors Snowflake se termine en $\Oh{\log n}$ tours avec une forte probabilité. 
\label{lemma:centrallimit}
\end{lemma}
\begin{proof}
%O The results follows from central limit theorem, wherein for $\alpha = \floor{k/2} + 1$, the expected bias in the network after sampling will be $\Oh{\sqrt{n}}$. An adversary smaller than this bias will be unable to keep the network in a fully-bivalent state for more than a constant number of rounds. The logarithmic factor remains from the mixing time lower bound. 
Le résultat procède du théorème central limite, où pour $\alpha = \floor{k/2} + 1$, la polarisation % HOULA
attendue dans le réseau après échantillonnage sera $\Oh{\sqrt{n}}$. Un adversaire plus petit que ce résultat sera incapable de conserver le réseau dans un état totalement bivalent pendant plus qu'un nombre constant de tours. Le facteur logarithmique provient de la limite inférieure du temps de mélange. % HOULA
\end{proof}
